\fred{A Unified Treatment of Tractabilty of Approximation Problems}

### 2023-06-25 GeneralizedTractability_230607.tex
Updates from Overleaf

@@

\title{\fred{A Unified Treatment of Tractabilty of Approximation Problems}\\ \peter{[PETER: In view of the papers of Henryk and Michael, maybe we should think of a better title?]}}

\author[1]{Fred J. Hickernell}
\ead{hickernell@iit.edu}
\author[2]{Peter Kritzer}
\author[1]{Onyekachi Osisiogu}
\date{\today}
\affiliation[1]{organization={Department of Applied Mathematics, Illinois Institute of Technology},
addressline={10 West 32${}^{\text{nd}}$ Street, Room 220},
postcode={60616},
city={Chicago, IL},
country={United States}}

@@

@@

The information complexity of the problem $\SOL_d$ is given by the function $\comp:(0,\infty) \times \naturals \to \natzero$, where $\comp(\varepsilon,d)$ is the number of linear functionals required by the best algorithm to satisfy the error criterion \eqref{eq:errcrit}.
By definition, the complexity is non-decreasing  as $\varepsilon$ tends to zero.  We also expect it to be non-decreasing as $d$ tends to infinity, since $d$ is typically the number of variables for the functions in $\cf_d$.

For this case where arbitrary linear functionals are allowed, the optimal solution operator is
\[
\APP_d(f,\varepsilon) = \sum_{i=1}^n \SOL(u_i) \ip[\cf_d]{f}{u_i},
\]
and the information complexity of our linear problem is
\begin{equation}\label{eq:comp_def}
\comp(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\}, \qquad \comp:(0,\infty) \times \naturals \to \natzero.
\end{equation}

The question of tractability is one of determining how fast the complexity increases as $\varepsilon$ tends to zero and/or as $d$ tends to infinity.  We want to bound the complexity in terms of a simple function, $T$, of $\varepsilon^{-1}$ and $d$ and identify simple conditions on the singular values that are equivalent to this bound holding.

In this article, we investigate a \emph{generalized tractability function}, denoted by $T$:
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty).
\end{equation}
The conditions we require $T$ to satisfy are described in  Section \ref{sec:not}, however, the basic idea is that we we define our approximation problem to be tractable if 
$\comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},d,\vp)$ for some constant, $C_{\vp}$, depending only on the parameter $\vp$.  In practice, we expect $T(\varepsilon^{-1},d,\vp)$ to increase mildly with increasing $\varepsilon^{-1}$ and/or $d$, however most of our theorems do not require such an assumption.

\Cref{thm_main_strong_tract2,thm_main_tract2,thm_main_rest_strong_tract,thm_main_rest_tract} all take the form
\begin{multline}
  \exists C_{\vp} \text{ with }
    \comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},\clubsuit,\vp) \quad \forall (\varepsilon^{-1},d) \in \Omega \\
    \iff   \exists L_{\vp} \text{ with }
    \sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,\clubsuit,\vp) \rceil}^{\spadesuit(d)} \frac{1}{T(\lambda^{-1}_{i,d},\clubsuit,\vp)}< \infty, \\
     \clubsuit \in \{1,d\}, \ \varheartsuit(d) := \inf \{\varepsilon : (\varepsilon^{-1},d) \in \Omega \}, \ 
     \spadesuit(d) := \begin{cases}
        \min\{n \in \natzero : \lambda_{n+1,d} \le \varheartsuit(d) \}, &   \varheartsuit(d) > 0, \\
        \infty, & \varheartsuit(d) = 0.
        \end{cases}
\end{multline}
\Cref{thm:subhT} is of similar, but not equivalent, form.

@@

In Section \ref{sec:spt}, we give the definitions of the considered tractability notions, in particular (strong) tractability. We consider the strong tractability when the information complexity is independent of the dimension, $d$. We then turn to tractability  when the information complexity is polynomial in $d$ and $\varepsilon^{-1}$. In Section \ref{sec:examples} we provide examples of strong(tractability) functions for both the algebraic and exponential cases. We also consider other notions of tractability such as quasi-polynomial tractability, and a generalized notion of weak tractability. In the case of the generalized weak tractability, we give the necessary and sufficient conditions on the notion of exponential tractability in Section \ref{sec:subh}.

@@

\subsection{Notation and fundamental definitions} \label{sec:not}
Let $T$ be a function given in \eqref{eq:Tspec}.
A tractability function, $T$ is a function of simple form that provides an upper bound on the computational complexity of a problem.  Some examples of tractability functions commonly considered are provided in Table \ref{tab:commonT}.

\begin{table} 
    \caption{Common forms of the tractability function, $T$}
{\small
\begin{equation*}
	\begin{array}{r@{\quad}c@{\quad}cc}
		\text{Tractability type} & T(\varepsilon^{-1},d,\vp)
		& T(0,d,\vp)\\
		\toprule
		\text{strong polynomial algebraic} & \max(1,\varepsilon^{-p}) & 1\\
		\text{polynomial algebraic} & \max(1,\varepsilon^{-p})d^{q} & d^{q}\\
		\text{strong polynomial exponential} &  [\max(1,\log(1 + \varepsilon^{-1})]^p & 1\\
		\text{polynomial exponential } &
		[\max(1,\log(1 + \varepsilon^{-1}))]^p  d^{q} & d^{q} \\
        \text{quasi-polynomial algebraic} & 
        \exp\{p(1+\log(\max(1,\varepsilon^{-1}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}\\
        \text{quasi-polynomial exponential} & 
        \exp\{p(1+\log(\max\{1,\log(1+\varepsilon^{-1})\}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}
	\end{array}
\end{equation*}}
\label{tab:commonT}
\end{table}

\begin{definition}
    A problem is defined as $T$-tractable with parameter $\vp$ iff there exists a positive constant $C_{\vp}$, which is independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},d,\vp) \qquad \forall \varepsilon >0, \ d \in \mathbb{N}.
\end{equation}
A problem is \emph{strongly}
$T$-tractable with parameter $\vp$ iff the complexity is independent of the dimension of the problem, that is, there exists a positive constant $C_{\vp}$, again independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:strong_tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},1,\vp) \qquad \forall \varepsilon >0 , \ d \in \mathbb{N}.
\end{equation}
We allow $\vp$ to be a scalar or vector as the situation dictates.
\end{definition}

For this definition of tractability to make sense we assume that
\begin{subequations} \label{eq:Tconditions}
\begin{equation} \label{eq:Tnondecreasing}
	T \text{ is non-decreasing in all variables,}
\end{equation}
which implies that the problem is expected to be no easier by decreasing the tolerance, $\varepsilon$, or increasing the dimension, $d$. (This is why we make $T$ a function of $\varepsilon^{-1}$ rather than of $\varepsilon$.) Furthermore, increasing $\vp$ does not decrease $T(\cdot, \cdot, \vp)$.  Since there are an infinite number of positive eigenvalues, it also makes sense to assume that
\begin{equation}
	\lim_{\varepsilon \to 0} T(\varepsilon^{-1},d,\vp) = \infty \qquad \forall d \in \naturals, \ \vp \in [0,\infty)^s.
\end{equation}
Since $T(\cdot,d,\vp)$ is non-decreasing, we may define the following limit, which we assume will always be positive:
\begin{equation}\label{eq:T0}	T(0,d,\vp):=\lim_{\varepsilon\to\infty}T(\varepsilon^{-1},d,\vp) = \inf_\varepsilon T(\varepsilon^{-1},d,\vp) \ge T(0,1,\vzero) > 0.
\end{equation}
\end{subequations}
Finally, there is a technical assumption required for our analysis.  There exists a $K_{\vp,\tau}$ depending on $\vp$ and $\tau$, but  independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:ptauassume}
	[T(\varepsilon^{-1},d,\vp)]^\tau \le K_{\vp,\tau} T(\varepsilon^{-1},d,\tau \vp),   \quad \forall \varepsilon \in (0,\infty), \ d \in \naturals, \ \vp\in[0,\infty)^s, \ \tau\in [1,\infty).
\end{equation}

@@

Since the $\lambda_{i,d}$ are non-increasing in $i$, it follows that the $T(\lambda_{i,d}^{-1},\clubsuit,\vp)$ are non-decreasing in $i$ for any $\clubsuit\in\naturals$. In particular, for  any $N\in\naturals$ with $n\ge N$, we have
\begin{align*}
    \lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{N,d}
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\clubsuit,\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},\clubsuit,\vp)} \le \cdots \le \frac{1}{T(\lambda_{N, d}^{-1},\clubsuit,\vp)} \\
     \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\clubsuit,\vp) }
    \le &\frac{1}{n-N+1} \sum_{i=N}^n  \frac{1}{T(\lambda_{i, d}^{-1},\clubsuit,\vp) }
    \le \frac{1}{n-N+1} \sum_{i=N}^\infty  \frac{1}{T(\lambda_{i, d}^{-1},\clubsuit,\vp)}.
\end{align*}
Thus, we can conclude from the previous line that
\begin{multline} \label{eq:implicate_a}
    n - N +1 \ge T(\varepsilon^{-1},\clubsuit,\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\clubsuit,\vp)} \\
   \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},\clubsuit,\vp)} \le
   \frac{1}{n-N+1} \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\clubsuit,\vp) } \le \frac{1}{T(\varepsilon^{-1},\clubsuit,\vp)} \\ \forall \clubsuit, n, N \in \naturals.
\end{multline}
We note that the left side of the implication implies $n \ge N$ automatically since $n$ is an integer  $n - N +1$ is assumed to be positive.

Moreover, given that $T$ is increasing in its arguments, we have the following equivalent expression for the complexity via \eqref{eq:comp_def}:
\begin{align} %\label{eq:comp_equiv}
	\nonumber
	\comp(\varepsilon, d) & = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\} \\
	\nonumber
	& = \min \biggl\{n \in \natzero : \frac{1}{T(\lambda_{n+1,d}^{-1},\clubsuit,\vp)}\le \frac{1}{T(\varepsilon^{-1},\clubsuit,\vp) }\biggr\} \qquad  \forall \clubsuit\in \naturals\\
	\nonumber
	& \le  \min \biggl\{n \in \natzero : n \ge N - 1 + T(\varepsilon^{-1},\clubsuit,\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\clubsuit,\vp)} \biggr\} \;  \forall  \clubsuit, N \in \naturals, \quad \text{by \eqref{eq:implicate_a}} \\ 
	& \le  T(\varepsilon^{-1},\clubsuit,\vp) \Biggl [\frac{N}{T(\varepsilon^{-1},\clubsuit,\vp)}  +  \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\clubsuit,\vp)} \Biggr] \quad \forall  \clubsuit, N \in \naturals. \label{eq:compUP}
\end{align}

Now we take this upper bound further, specializing to the case of $\clubsuit=1$ and $N=L_{\vp}$:
\begin{align*}
       \comp(\varepsilon,d)
       & \le T(\varepsilon^{-1},1, \vp) \Biggl [ \frac{L_{\vp}}{T(\varepsilon^{-1},1, \vp)}  + \underbrace{\sum_{i=L_{\vp}}^\infty \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)}}_{\le S_{\vp} <\infty\ \text{by \eqref{eq:strong_tractiff3}} }
        \Biggr] \\
       & \le T(\varepsilon^{-1},1,\vp) \underbrace{\left[\frac{L_{\vp}}{T(0,1, \vp)} + S_{\vp} \right]}_{=:C_{\vp}}
       \qquad \text{by \eqref{eq:Tnondecreasing}}\\
       & =  C_{\vp}T(\varepsilon^{-1},1,\vp).
%       & \le \underbrace{2 C_{1,\vp}}_{\text{since } T(\varepsilon^{-1},1,\vp) \ge 1} T(\varepsilon^{-1},1,\vp)
%       = \underbrace{C_{\vp}}_{:=2 C_{1,\vp}}T(\varepsilon^{-1},1,\vp).
\end{align*}
This means that we have strong $T$-tractability via \eqref{eq:strong_tract_def}, and verifies the sufficiency of \eqref{eq:strong_tractiff3}.

@@

For $\varepsilon$ taking on all positive values, $i(\varepsilon,\vp)$ takes on all values greater than or equal to $i(\infty,\vp)$, so
\[
i\le C_{\vp} T(\lambda_{i,d}^{-1},1,\vp)+1, \qquad \forall i \ge i(\infty,\vp), \ d \in \naturals.
\]
This implies via our technical assumption \eqref{eq:ptauassume} that
\begin{gather*}
 K_{\vp,\tau}\,T (\lambda_{i,d}^{-1},1,\tau \vp) \ge
 [T(\lambda_{i,d}^{-1},1, \vp)]^\tau
 \ge
  \left[\frac{(i-1)}{C_{\vp}}\right]^\tau \qquad \forall i \ge i(\infty,\vp)), \ d \in \naturals, \\
 \frac{1}{T (\lambda_{i,d}^{-1},1,\tau \vp)} \le
\frac{K_{\vp,\tau}\, C_{\vp}^\tau}{(i-1)^\tau}\qquad \forall i \ge \max\{i(\infty,\vp),2\}), \ d \in \naturals .
\end{gather*}
Summing both sides of this last inequality over $i$ from $i(\infty,\vp) +1 \ge 2$ to $\vinfty$ yields
\begin{equation*}
\sup_{d\in\naturals} \sum_{i=i(\infty,\vp) +1 }^\infty \frac{1}{T (\lambda_{i,d}^{-1},1, \tau \vp)}
 \le  K_{\vp,\tau}\, C_{\vp}^\tau
\sum_{i=2}^\infty \frac{1}{(i-1)^\tau} \\
 \le K_{\vp,\tau}\, C_{\vp}^\tau
\zeta (\tau)  < \infty,
\end{equation*}
where $\zeta$ denotes the Riemann zeta function.
This yields \eqref{eq:strong_tractiff3} with $\vp$ replaced by $\vp' = \tau \vp$ and $L_{\vp'} = i(\infty,\vp) +1$, and so we see the necessity of $\eqref{eq:strong_tractiff3}$.

@@

\begin{proof}
    \textbf{Sufficient Condition:}\\
Suppose that \eqref{eq:tractiff4} holds for some $\vp \ge 0$. 
By the argument in the proof of the sufficient condition for Theorem \ref{thm_main_strong_tract2}, it follows from \eqref{eq:compUP} with $\clubsuit=d$ and $N = \lceil L_{\vp}\, T(0,d,\vp)\rceil$ that 
\begin{align*}
       \comp(\varepsilon,d)
       & \le T(\varepsilon^{-1},d,\vp) \Biggl [ 
       \frac{\lceil L_{\vp}\, T(0,d,\vp)\rceil}{T(\varepsilon^{-1},d,\vp)}  + 
       \underbrace{\sum_{i=\lceil L_{\vp}\, T(0,d,\vp)\rceil}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:tractiff4}}}
       \Biggr ] 
        \quad \forall \varepsilon > 0, \ d  \in \naturals\\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp} T(0,d,\vp)+1}{T(\varepsilon^{-1},d, \vp)} +  S_{\vp} \right]
        \qquad \forall \varepsilon > 0, \ d  \in \naturals\\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp} T(0,d,\vp)}{T(\varepsilon^{-1},d, \vp)}
       + \frac{1}{T(0,1, \vp)} + S_{\vp}\right]
       \qquad \text{by \eqref{eq:Tnondecreasing}}   \\
       & \le T(\varepsilon^{-1},d,\vp) 
       \underbrace{\left [  L_{\vp} + \frac{1}{T(0,1, \vp)} + S_{\vp} \right]}_{=:C_{\vp}}
       \qquad \text{by \eqref{eq:Tnondecreasing}}  
\end{align*}
It follows that we have $T$-tractability, which shows sufficiency of \eqref{eq:tractiff4}.

@@

The (strong) polynomial tractability for both the algebraic case and the exponential case will be first considered as follows: 
\begin{example}[Strong algebraic polynomial tractability]

>>>>

%certain important notations for key symbols
\newcommand{\thed}{\clubsuit}
\newcommand{\theM}{\varheartsuit}
\newcommand{theUB}{\spadesuit}

@@

\title{\fred{A Unified Treatment of Tractabilty of Approximation Problems}}

\author[1]{Fred J. Hickernell}
\ead{hickernell@iit.edu}
\author[2]{Peter Kritzer}
\ead{peter.kritzer@oeaw.ac.at}
\author[1]{Onyekachi Osisiogu}
\date{\today}
\affiliation[1]{organization={Department of Applied Mathematics, Illinois Institute of Technology},
addressline={10 West 32${}^{\text{nd}}$ Street, Room 220},
postcode={60616},
city={Chicago, IL},
country={United States}}
\affiliation[2]{organization={Johann Radon Institute for Computational and Applied Mathematics (RICAM), Austrian Academy of Sciences},
addressline={Altenbergerstr. 69},
postcode={4040},
city={Linz},
country={Austria}}

@@

TODO: We should check for consistency of the use of the $\forall$ symbol in displayed formulas. Sometimes we use it, sometimes we do not, sometimes it is preceded by a comma, sometimes not, etc.

@@

The information complexity of the problem $\SOL_d$ is given by the function $\comp:(0,\infty) \times \naturals \to \natzero$, where $\comp(\varepsilon,d)$ is the number of linear functionals required by the best \peter{admissible} algorithm to satisfy the error criterion \eqref{eq:errcrit}.
By definition, the complexity is non-decreasing  as $\varepsilon$ tends to zero.  We also expect it to be non-decreasing as $d$ tends to infinity, since $d$ is typically the number of variables for the functions in $\cf_d$.

For this case where arbitrary linear functionals are allowed, the optimal solution operator is
\[
\APP_d(f,\varepsilon) = \sum_{i=1}^n \peter{\SOL_d(u_{i,d})} \ip[\cf_d]{f}{u_i},
\]
and the information complexity of our linear problem is
\begin{equation}\label{eq:comp_def}
\comp(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\}.
\end{equation}

The question of tractability is one of determining how fast the \peter{information} complexity increases as $\varepsilon$ tends to zero and/or as $d$ tends to infinity.  We want to bound the complexity in terms of a simple function, $T$, of $\varepsilon^{-1}$ and $d$ and identify simple conditions on the singular values that are equivalent to this bound holding. \peter{[PETER: Maybe this would be a good place to refer to the earlier work of Michael and Henryk? I will take care of it... .]}

In this article, we investigate a \emph{generalized tractability function}, denoted by $T$:
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty).
\end{equation}
The conditions we require $T$ to satisfy are described in  Section \ref{sec:not}, however, the basic idea is that we define our approximation problem to be tractable if 
$\comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},d,\vp)$ for some constant, $C_{\vp}$, depending only on the parameter $\vp$.  In practice, we expect $T(\varepsilon^{-1},d,\vp)$ to increase sub-exponentially  with increasing $\varepsilon^{-1}$ and/or $d$, however most of our theorems do not require such an assumption.

\Cref{thm_main_strong_tract2,thm_main_tract2,thm_main_rest_strong_tract,thm_main_rest_tract} all take the form

\peter{[PETER: Not meaning to sound too negative, but I am not sure I like the notation with the symbols below. Is it possible to use something different instead?]}
\fred{Fred: I have now given them names above in lines 18-20, so we can experiment with different symbols or letters.}
\begin{multline}
  \exists C_{\vp} \text{ with }
    \comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},\thed,\vp) \quad \forall (\varepsilon^{-1},d) \in \Omega \\
    \iff   \exists L_{\vp} \text{ with }
    \sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,\thed,\vp) \rceil}^{\theUB(d)} \frac{1}{T(\lambda^{-1}_{i,d},\thed,\vp)}< \infty, \\
     \thed \in \{1,d\}, \ \theM(d) := \inf \{\varepsilon : (\varepsilon^{-1},d) \in \Omega \}, \ 
     \theUB(d) := \begin{cases}
        \min\{n \in \natzero : \lambda_{n+1,d} \le \theM(d) \}, &   \theM(d) > 0, \\
        \infty, & \theM(d) = 0.
        \end{cases}
\end{multline}
\Cref{thm:subhT} is of similar, but not equivalent, form.

@@

In Section \ref{sec:spt}, we give the definitions of the considered tractability notions, in particular (strong) tractability. We consider strong tractability when the information complexity is independent of the dimension, $d$. We then turn to tractability  when the information complexity is polynomial in $d$ and $\varepsilon^{-1}$. In Section \ref{sec:examples} we provide examples of strong(tractability) functions for both the algebraic and exponential cases. We also consider other notions of tractability such as quasi-polynomial tractability, and a generalized notion of weak tractability. In the case of the generalized weak tractability, we give the necessary and sufficient conditions on the notion of exponential tractability in Section \ref{sec:subh}.

@@

\subsection{Notation and fundamental definitions} \label{sec:not}
Let $T$ be a function given in \eqref{eq:Tspec}.
A tractability function $T$ is a function of simple form that provides an upper bound on the \peter{information} complexity of a problem.  Some examples of tractability functions commonly considered are provided in Table \ref{tab:commonT}.

\begin{table} 
    \caption{Common forms of the tractability function, $T$. \peter{[PETER: I changed the notation below, using ``algebraic" and ``exponential" first, to be in line with the literature.]}}
{\small
\begin{equation*}
	\begin{array}{r@{\quad}c@{\quad}cc}
		\text{Tractability type} & T(\varepsilon^{-1},d,\vp)
		& T(0,d,\vp)\\
		\toprule
		\text{\peter{algebraic strong polynomial}} & \max(1,\varepsilon^{-p}) & 1\\
		\text{\peter{algebraic polynomial}} & \max(1,\varepsilon^{-p})d^{q} & d^{q}\\
		\text{\peter{exponential strong polynomial}} &  [\max(1,\log(1 + \varepsilon^{-1})]^p & 1\\
		\text{\peter{exponential polynomial}} &
		[\max(1,\log(1 + \varepsilon^{-1}))]^p  d^{q} & d^{q} \\
        \text{\peter{algebraic quasi-polynomial}} & 
        \exp\{p(1+\log(\max(1,\varepsilon^{-1}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}\\
        \text{\peter{exponential quasi-polynomial}} & 
        \exp\{p(1+\log(\max\{1,\log(1+\varepsilon^{-1})\}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}
	\end{array}
\end{equation*}}
\label{tab:commonT}
\end{table}

\begin{definition}
    A problem is defined as $T$-tractable with parameter $\vp$ iff there exists a positive constant $C_{\vp}$, which is independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},d,\vp) \qquad \forall \varepsilon >0, \ d \in \mathbb{N}.
\end{equation}
A problem is \emph{strongly}
$T$-tractable with parameter $\vp$ iff the \peter{information} complexity is independent of the dimension of the problem, that is, there exists a positive constant $C_{\vp}$, again independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:strong_tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},1,\vp) \qquad \forall \varepsilon >0 , \ d \in \mathbb{N}.
\end{equation}
We allow $\vp$ to be a scalar or vector as the situation dictates.
\end{definition}

For this definition of tractability to make sense we assume that
\begin{subequations} \label{eq:Tconditions}
\begin{equation} \label{eq:Tnondecreasing}
	T \text{ is non-decreasing in all variables,}
\end{equation}
which implies that the problem is expected to be no easier by decreasing the tolerance, $\varepsilon$, or increasing the dimension, $d$. (This is why we make $T$ a function of $\varepsilon^{-1}$ rather than of $\varepsilon$.) Furthermore, increasing $\vp$ does not decrease $T(\cdot, \cdot, \vp)$.  Since there are an infinite number of positive eigenvalues, it also makes sense to assume that
\begin{equation}\label{eq:Tlim}
	\lim_{\varepsilon \to 0} T(\varepsilon^{-1},d,\vp) = \infty \qquad \forall d \in \naturals, \ \vp \in [0,\infty)^s.
\end{equation}
Since $T(\cdot,d,\vp)$ is non-decreasing, we may define the following limit, which we assume will always be positive:
\begin{equation}\label{eq:T0}	T(0,d,\vp):=\lim_{\varepsilon\to\infty}T(\varepsilon^{-1},d,\vp) = \inf_\varepsilon T(\varepsilon^{-1},d,\vp) \ge T(0,1,\vzero) > 0.
\end{equation}
\end{subequations}
Finally, there is a technical assumption required for our analysis.  There exists a $K_{\vp,\tau}$ depending on $\vp$ and $\tau$, but  independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:ptauassume}
	[T(\varepsilon^{-1},d,\vp)]^\tau \le K_{\vp,\tau} T(\varepsilon^{-1},d,\tau \vp),   \quad \forall \varepsilon \in (0,\infty), \ d \in \naturals, \ \vp\in[0,\infty)^s, \ \tau\in [1,\infty).
\end{equation}

@@

Since the $\lambda_{i,d}$ are non-increasing in $i$, it follows that the $T(\lambda_{i,d}^{-1},\thed,\vp)$ are non-decreasing in $i$ for any $\thed\in\naturals$. In particular, for  any $N\in\naturals$ with $n\ge N$, we have
\begin{align*}
    \lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{N,d}
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},\thed,\vp)} \le \cdots \le \frac{1}{T(\lambda_{N, d}^{-1},\thed,\vp)} \\
     \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp) }
    \le &\frac{1}{n-N+1} \sum_{i=N}^n  \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp) }
    \le \frac{1}{n-N+1} \sum_{i=N}^\infty  \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp)}.
\end{align*}
Thus, we can conclude from the previous line that
\begin{multline} \label{eq:implicate_a}
    n - N +1 \ge T(\varepsilon^{-1},\thed,\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp)} \\
   \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp)} \le
   \frac{1}{n-N+1} \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp) } \le \frac{1}{T(\varepsilon^{-1},\thed,\vp)} \\ \forall \thed, n, N \in \naturals.
\end{multline}
We note that the left side of the implication implies $n \ge N$ automatically since $n$ is an integer  $n - N +1$ is assumed to be positive.

Moreover, given that $T$ is increasing in its arguments, we have the following equivalent expression for the complexity via \eqref{eq:comp_def}:
\begin{align} %\label{eq:comp_equiv}
	\nonumber
	\comp(\varepsilon, d) & = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\} \\
	\nonumber
	& = \min \biggl\{n \in \natzero : \frac{1}{T(\lambda_{n+1,d}^{-1},\thed,\vp)}\le \frac{1}{T(\varepsilon^{-1},\thed,\vp) }\biggr\} \qquad  \forall \thed\in \naturals\\
	\nonumber
	& \le  \min \biggl\{n \in \natzero : n \ge N - 1 + T(\varepsilon^{-1},\thed,\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp)} \biggr\} \;  \forall  \thed, N \in \naturals, \quad \text{by \eqref{eq:implicate_a}} \\ 
	& \le  T(\varepsilon^{-1},\thed,\vp) \Biggl [\frac{N}{T(\varepsilon^{-1},\thed,\vp)}  +  \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp)} \Biggr] \quad \forall  \thed, N \in \naturals. \label{eq:compUP}
\end{align}

Now we take this upper bound further, specializing to the case of $\thed=1$ and $N=L_{\vp}$:
\begin{align*}
       \comp(\varepsilon,d)
       & \le T(\varepsilon^{-1},1, \vp) \Biggl [ \frac{L_{\vp}}{T(\varepsilon^{-1},1, \vp)}  + \underbrace{\sum_{i=L_{\vp}}^\infty \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)}}_{\le S_{\vp} <\infty\ \text{by \eqref{eq:strong_tractiff3}} }
        \Biggr] \\
       & \le T(\varepsilon^{-1},1,\vp) \underbrace{\left[\frac{L_{\vp}}{T(0,1, \vp)} + S_{\vp} \right]}_{=:C_{\vp}}
       \qquad \text{by \eqref{eq:Tnondecreasing}}\\
       & =  C_{\vp}T(\varepsilon^{-1},1,\vp).
%       & \le \underbrace{2 C_{1,\vp}}_{\text{since } T(\varepsilon^{-1},1,\vp) \ge 1} T(\varepsilon^{-1},1,\vp)
%       = \underbrace{C_{\vp}}_{:=2 C_{1,\vp}}T(\varepsilon^{-1},1,\vp).
\end{align*}
This means that we have strong $T$-tractability via \eqref{eq:strong_tract_def}, and verifies the sufficiency of \eqref{eq:strong_tractiff3}.

@@

For $\varepsilon$ taking on all positive values, $i(\varepsilon,\vp)$ takes on, \peter{by \eqref{eq:Tlim},} all values greater than or equal to $i(\infty,\vp)$, so
\[
i\le C_{\vp} T(\lambda_{i,d}^{-1},1,\vp)+1, \qquad \forall i \ge i(\infty,\vp), \ d \in \naturals.
\]
This implies via our technical assumption \eqref{eq:ptauassume} that, \peter{for any $\tau>1$,}
\begin{gather*}
 K_{\vp,\tau}\,T (\lambda_{i,d}^{-1},1,\tau \vp) \ge
 [T(\lambda_{i,d}^{-1},1, \vp)]^\tau
 \ge
  \left[\frac{(i-1)}{C_{\vp}}\right]^\tau \qquad \forall i \ge i(\infty,\vp)), \ d \in \naturals, \\
 \frac{1}{T (\lambda_{i,d}^{-1},1,\tau \vp)} \le
\frac{K_{\vp,\tau}\, C_{\vp}^\tau}{(i-1)^\tau}\qquad \forall i \ge \max\{i(\infty,\vp),2\}), \ d \in \naturals .
\end{gather*}
Summing both sides of this last inequality over $i$ from $i(\infty,\vp) +1 \ge 2$ to $\vinfty$ yields
\begin{equation*}
\sup_{d\in\naturals} \sum_{i=i(\infty,\vp) +1 }^\infty \frac{1}{T (\lambda_{i,d}^{-1},1, \tau \vp)}
 \le  K_{\vp,\tau}\, C_{\vp}^\tau
\sum_{i=2}^\infty \frac{1}{(i-1)^\tau} \\
 \le K_{\vp,\tau}\, C_{\vp}^\tau
\zeta (\tau)  < \infty,
\end{equation*}
where $\zeta$ denotes the Riemann zeta function.
This yields \eqref{eq:strong_tractiff3} with $\vp$ replaced by $\vp' = \tau \vp$ and $L_{\vp'} = i(\infty,\vp) +1$, and so we see the necessity of $\eqref{eq:strong_tractiff3}$.

@@

\begin{proof}
    \textbf{Sufficient condition:}\\
Suppose that \eqref{eq:tractiff4} holds for some $\vp \ge 0$. 
By the argument in the proof of the sufficient condition for Theorem \ref{thm_main_strong_tract2}, it follows from \eqref{eq:compUP} with $\thed=d$ and $N = \lceil L_{\vp}\, T(0,d,\vp)\rceil$ that 
\begin{align*}
       \comp(\varepsilon,d)
       & \le T(\varepsilon^{-1},d,\vp) \Biggl [ 
       \frac{\lceil L_{\vp}\, T(0,d,\vp)\rceil}{T(\varepsilon^{-1},d,\vp)}  + 
       \underbrace{\sum_{i=\lceil L_{\vp}\, T(0,d,\vp)\rceil}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:tractiff4}}}
       \Biggr ] 
        \quad \forall \varepsilon > 0, \ d  \in \naturals\\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp} T(0,d,\vp)+1}{T(\varepsilon^{-1},d, \vp)} +  S_{\vp} \right]
        \qquad \forall \varepsilon > 0, \ d  \in \naturals\\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp} T(0,d,\vp)}{T(\varepsilon^{-1},d, \vp)}
       + \frac{1}{T(0,1, \vp)} + S_{\vp}\right]
       \qquad \text{by \eqref{eq:Tnondecreasing}}   \\
       & \le T(\varepsilon^{-1},d,\vp) 
       \underbrace{\left [  L_{\vp} + \frac{1}{T(0,1, \vp)} + S_{\vp} \right]}_{=:C_{\vp}}
       \qquad \text{by \eqref{eq:Tnondecreasing}} .
\end{align*}
It follows that we have $T$-tractability, which shows sufficiency of \eqref{eq:tractiff4}.

@@

The (strong) polynomial tractability for both the algebraic case and the exponential case will be first considered as follows:
